---
title: "Homework 5"
author: "Irene Yang"
date: "3/14/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Prep}

#Load all packages
library(phyloseq)
packageVersion("phyloseq")
library(ggplot2) 
packageVersion("ggplot2")
library(RColorBrewer)
packageVersion("RColorBrewer")

# Load HMP data
# You will need to download the data from Canvas
# You will need to change your path

load("~/Desktop/N741/2018Week8/HMPv35.Rdata")

HMPv35

```

####Problem 1

Subset the HMPv35 object to obtain only the samples from Tongue_dorsum. Call this new object HMPv35sub2

```{r}

# Subset

sub <- get_variable(HMPv35, "HMPbodysubsite") %in% c("Tongue_dorsum")

sample_data(HMPv35)$sub <- factor(sub)

HMPv35sub2 <- prune_samples(sample_data(HMPv35)$sub == TRUE, HMPv35)

summary(sample_data(HMPv35sub2))

HMPv35sub2

```

####Problem 2

Produce the geometric box plot of diversity measures for your object, HMPv35sub2

```{r Diversity}

# Changing the background to my favorite
theme_set(theme_bw())

# The following commands plot the different diversity measures for the two different body sites.
alpha_meas = c("Observed", "Chao1", "ACE", "Shannon", "Simpson", "InvSimpson")
(p <- plot_richness(HMPv35sub2, "HMPbodysubsite", measures=alpha_meas))

# Boxplot to display diversity

p + geom_boxplot(data=p$data, aes(x=HMPbodysubsite, y=value, color=NULL), alpha=0.1)

```

####Problem 3

You see what taxa are most prevalent in your subset, HMPv35sub2

```{r Looking at phyla}

ntaxa(HMPv35sub2)

par(mar = c(10, 4, 4, 2) + 0.1) # make more room on bottom margin

#Looking at top 20

N <- 20

barplot(sort(taxa_sums(HMPv35sub2), TRUE)[1:N]/nsamples(HMPv35sub2), las=2)

```

####Problem 4

Using your HMPv35sub2 object, throw the rare taxa out of that object, then reduce to only taxa in the phylum Bacteroidetes. Call this new object HMPv35sub2frbac 

```{r Filtering data}

#The next step filters out taxa with low occurrence throughout all samples

HMPv35subsub = filter_taxa(HMPv35sub2, function(x) sum(x > 3) > (0.2*length(x)), TRUE)

# The next step filters out all taxa that occur in less than .01% of samples

HMPv35subr  <- transform_sample_counts(HMPv35subsub, function(x) x / sum(x) )
HMPv35subfr <- filter_taxa(HMPv35subr, function(x) mean(x) > 1e-5, TRUE)

HMPv35subfr

#Finally subset to only bacteria in the phylum Bacteroidetes

HMPv35sub2frbac = subset_taxa(HMPv35subfr, Phylum=="Bacteroidetes")
HMPv35sub2frbac

```

####Problem 5
Using your HMPv35sub2frbac object, what is the distribution of the top 20 OTU's?

```{r}

# Calculate number of taxa
ntaxa(HMPv35sub2frbac)

# Distribution of top 20 OTU's
par(mar = c(10, 4, 4, 2) + 0.1) # make more room on bottom margin
N <- 20
barplot(sort(taxa_sums(HMPv35sub2frbac), TRUE)[1:N]/nsamples(HMPv35sub2frbac), las=2)

```

####Problem 6

Plot sample abundance by body site for your object HMPv35sub2frbac

```{r}

title = "plot_bar; by site; Bacteroidetes only"
plot_bar(HMPv35sub2frbac, "HMPbodysubsite", "Abundance", title=title)

```

####Problem 7

You try it with your HMPv35sub2frbac object

```{r}

# Differentiating by Family

plot_bar(HMPv35sub2frbac, "HMPbodysubsite", "Abundance", "Family", title=title)
#Place your code for your plot with family here

```

We can use the techniques of network science to illustrate how similar or distant samples are.

Using our filtered object with all phyla, we use the `plot_net` function to illustrate the "proximity" (or similarity) of samples, while denoting site and the sex of the participant.

```{r}
plot_net(HMPv35subfr, maxdist = 0.3, color = "HMPbodysubsite", shape="sex")
```

Let's redo switching which variable is colored and which variable is differentiated by shape.

```{r}
plot_net(HMPv35subfr, maxdist = 0.3, shape = "HMPbodysubsite", color="sex")
```
####Problem 8

Which method of display do you like best and why?

```{r}

#Place your answer here as another comment.

# Definitely like the second display better. Differentiating sex by color makes the graphic much easier to visualize and interpret.

```
